Indirect Interventions for Stabilizing Housing Markets

### 2022-04-26 outputs/paper/paper.Rmd
add summary statistics table

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

@@

# Data
[@citeR]

@@

# Enhancement

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(kableExtra)
library(here)
```

@@

# Data
We would like to investigate the relationship between changes in unemployment benefit generosity and the number of home foreclosures within the United States. The original analysis performed by [@citeHsu] is done in Stata. For the purposes of aiding reproducibility, we will be reproducing their scripts using `R` [@citeR], `dplyr` [@citeDplyr] and converting data filetypes with `haven` [@citeHaven]. Tables are generated using `kableExtra` [@citeKable].

As a reproduction, we will be utilising the same data as the original paper, as made available and obtained from openICPSR (CITATION). There are five separate datasets used within the paper:

- Attributes relating to unemployment benefits for each state for the years of 1991-2010, obtained from the Bureau of Labor Statistics, Bureau of Economic Analysis, Department of Labor, and Hirsch and Macpherson (2003).

- Factors of mortgage borrowers relating to potential of delinquency from the Survey of Income and Program Participation (SIPP) years 1991, 1992, 1993, 1996, 2000, 2004, 2008 and supplemented by the corresponding unemployment insurance program characteristics obtained from the Department of Labor.

- Household data of mortgage borrowers from the NLSY79 survey years 2010 and 2012 for foreclosure related variables.

- Characteristics of mortgages and housing relating to foreclosure risk by state from NLSY79 2010 and 2012.

- Home price growth, unemployment rates, and unemployment insurance data of each state from 2008 to 2013, obtained from Zillow, Bureau of Labor Statistics, and Department of Labor respectively.

```{r summary_stats}
# read in csvs for each dataset
econ_analysis <- read_csv(here::here("inputs/data/econ_analysis.csv"))
sipp_analysis <- read_csv(here::here("inputs/data/sipp_analysis.csv"))
nlsy_analysis <- read_csv(here::here("inputs/data/nlsy_analysis.csv"))
nlsy_state_variables <- read_csv(here::here("inputs/data/nlsy_state_variables.csv"))
hpi_analysis <- read_csv(here::here("inputs/data/hpi_analysis.csv"))

# replicate table 1 of original paper, get mean median and sd for all variables

# Panel A, state characteristics
panel_a <- econ_analysis %>%
  select(max_ben, wba_max_thousands, duration_max, max_ben_real, ln_max_ben, 
         max_ben_wages, ui_rr, neg_ui_rr_pct, max_ben_eb_euc, eb_euc_weeks,
         unemp_rate, ln_realgdp_percap, wages_state, cov) %>%
  pivot_longer(everything(), names_to = 'variable', values_to = 'value') %>%
  group_by(variable) %>%
  summarise(mean = mean(value, na.rm = TRUE), median = median(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE))

# Panel B, household characteristics, delinquency analysis
panel_b <- sipp_analysis %>%
  select(delinq_mort, evict, ltv_win, neg_equity, mortgage_ui, max_ben_indiv,
         layoff, earnings_total, assets_liquid, thhtnw, educ_less_hs, educ_hs,
         educ_somecol, educ_col, educ_grad) %>%
  pivot_longer(everything(), names_to = 'variable', values_to = 'value') %>%
  group_by(variable) %>%
  summarise(mean = mean(value, na.rm = TRUE), median = median(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE))

# Panel C, household characteristics, foreclosure analysis
panel_c <- nlsy_analysis %>%
  select(foreclosure, lost_home, layoff, earnings_lag, net_worth_lag, 
         ltv_win_lag, ltv_win2nd_lag, neg_equity_lag, neg_equity2nd_lag) %>%
  pivot_longer(everything(), names_to = 'variable', values_to = 'value') %>%
  group_by(variable) %>%
  summarise(mean = mean(value, na.rm = TRUE), median = median(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE))

# create table using kableExtra
all_panels <- bind_rows(panel_a, panel_b, panel_c)

all_panels %>% 
  kbl(booktabs = T, col.names = c('Variable', 'Mean', 'Median', 'SD'), digits = 2) %>%
  kable_styling(position = "center")
```

@@

# Enhancement
## Datasheet for Dataset
